### 2017-06-06 manuscript/Appendix1.Rmd
beginnings of appendix 1

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \linespread{1.25}
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A1-\arabic{figure}}  
\renewcommand\thetable{A1-\arabic{table}}  
\renewcommand\thesection{Section A1.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 1}} \\
A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
``Ecosystem and community resistance...'' \\
\emph{PeerJ}
\end{center}

\linenumbers

\section{Estimating ANPP}
We used a radiometer to nondestructively estimate aboveground net primary productivity.
Our approach relies on relating greenness in a plot to aboveground biomass.

\section{Additional Figures}