### 2017-06-12 manuscript/Appendix3.Rmd
adds stuff on hrv and extreme years

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A3-\arabic{figure}}
\renewcommand\theequation{A3.\arabic{equation}}
\renewcommand\thetable{A3-\arabic{table}}  
\renewcommand\thesection{Section A3.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 3}} \\
A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
``Ecosystem and community resistance...'' \\
\emph{PeerJ}
\end{center}

\section{Characterizing Extreme Precipitation Amounts}

Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
The data consist of 91 yearly records, which we assume are approximately normally distributed.
The R code below shows our procedure, and Fig. A3-1 shows the results.

```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
library(ggplot2)
weather    <- read.csv("../data/weather/ClimateIPM.csv")
mean_ppt   <- mean(weather$ppt1)
quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation

ggplot(weather, aes(x=ppt1))+
  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
  geom_line(stat="density", color="blue")+
  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
  ylab("Density")+
  xlab("Growing Season Precipitation (mm)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

# References