### 2017-10-31 manuscript/review_response_R1.Rmd
starts response letter

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \linespread{1}
  - \usepackage{rotating}
  - \usepackage[usenames]{color}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
  - \usepackage{mathtools}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{R1-\arabic{figure}}  
\renewcommand\thetable{R1-\arabic{table}}

\definecolor{ForestGreen}{rgb}{0.1,0.44,0.1} 
\definecolor{blue}{rgb}{0,0,0.7} 
\definecolor{red}{rgb}{0.8,0,0}

\newcommand{\new}{\textcolor{red}} 
\newcommand{\comment}{\textcolor{ForestGreen}}
\newcommand{\response}{\textcolor{blue}}

\begin{center}
\textbf{\Large{Response to Reviews}}
\end{center}

\begin{flushright}
Date: \today
\end{flushright}

Dr. Buckley,

Thank you for the opportunity to revise our manuscript.
All three reviewers agree our paper is interesting, but point out several areas where we could improve the analysis.
We were impressed with the very constructive reviews, and thank the reviewers for taking such care when reading our paper.
We hope our paper is now suitable for publication in \emph{PeerJ}.

In our point-by-point response below, the original reviews are in black and \response{our responses are in blue.} 
Likewise, changes made in the manuscript text are \response{in blue} for clarity.
In our responses, we refer to line numbers of the revised manuscript unless otherwise stated.

\bigskip{}

\normalsize 
\hrulefill

Manuscript number: 20739  
Title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe  
Author(s): Tredennick, A.T.; Kleinhesselink, A.R.; Taylor, J.B.;  Adler, P.B.

\hrulefill
\bigskip{} \bigskip{}

\textbf{Editors Comments for the Author(s)}

Please pay close attention to the comments from Reviewer 1 regarding the validity of the findings. Once you have responded to these comments in a point-by-point cover letter and using tracked changes in the manuscript document, I will likely send your manuscript out for re-review.

\response{We provide point-by-point responses to reviewer comments below. We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by }

\bigskip{}

\textbf{Reviewer: 1}