### 2017-12-10 manuscript/Appendix4.Rmd

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A4-\arabic{figure}}
\renewcommand\theequation{A4.\arabic{equation}}
\renewcommand\thetable{A4-\arabic{table}}  
\renewcommand\thesection{Section A4.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 4}} \\
A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

\section{Conducting our analysis with NDVI}

The first step of our analysis is to convert NDVI from field-based radiometric measurements to aboveground net primary productivity.
However, this conversion is not perfect and the associated uncertainty is sometimes high (Table A1-2).
Therefore, we also assessed ecosystem functional response across treatments using NDVI as the response variable rather than ANPP.
The model essentially the same as the ANPP model, except we use a Beta likelihood for NDVI since its values range from 0 to 1, not including true 0s or 1s (see Stan code, below).

Our NDVI-model results are very similar to the ANPP-model results (Figure A4-1).
The only difference is there is a higher probability that the drought slope offset is less than 1 (90\% probability).
However, these negative offsets do not result in significantly different slopes once applied to the control slope, just as in the ANPP-model.

```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the NDVI-soil moisture relationship. Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots. Probabilities (Pr </> 0 =) for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero. The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.", include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}

library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots

all_fit <- readRDS("../results/randcoefs_alltreatments_NDVIfit.RDS")

####
####  FUNCTION TO CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) ----
####
get_one_tailed <- function(values){
  above <- 1 - ecdf(values)(0)
  below <- ecdf(values)(0)
  max(above,below)
}

####
####  PLOT TREATMENT-LEVEL POSTERIOR DISTRIBUTIONS ----
####
mycols <- c("#009E73", "#D55E00", "#0072B2")
param_labels <- c("Control::Intercept",
                  "Drought::Intercept", 
                  "Irrigation::Intercept",
                  "Control::Slope", 
                  "Drought::Slope", 
                  "Irrigation::Slope")

betas <- data.frame(extract(all_fit, pars = 'beta')) %>% 
  mutate(iteration = row_number()) %>% 
  gather(param_name, estimate, starts_with('beta')) %>%
  mutate(param_name = factor(param_name, labels = param_labels)) %>%
  separate(param_name, c("Treatment", "Type"), "::")

treat_slopes <- betas %>%
  filter(Treatment != "Control") %>%
  mutate(Type = paste(Type,"Offset"))

slope_probs <- treat_slopes %>%
  group_by(Treatment, Type) %>%
  summarise(probs = round(get_one_tailed(estimate),2)) %>%
  mutate(prob_text = paste("Pr < 0 =", probs))
slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
slope_probs$ypos <- c(1.3,1.3,2.55,2.7)

treat_cols <- mycols[2:3]

ggplot(treat_slopes)+
  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
  geom_joy(stat="density", adjust=3, alpha=0.8, aes(x=estimate, y=Treatment, fill=Treatment, color=Treatment, height = ..density..))+
  geom_text(data = slope_probs, aes(x=xpos, y=ypos, label = prob_text), size = 3)+
  facet_wrap(~Type, scales = "free")+
  scale_y_discrete(labels = c("Drought","Irrigation"))+
  scale_fill_manual(values = treat_cols)+
  scale_color_manual(values = treat_cols)+
  guides(fill=FALSE, color=FALSE)+
  xlab("Coefficient Value")+
  ylab(NULL)+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
        panel.border     = element_blank(), 
        axis.line        = element_blank(),
        axis.ticks       = element_blank(),
        strip.background = element_blank(),
        strip.text       = element_text(face="bold"),
        panel.spacing    = unit(1, units = "cm"))
```

\newpage{}

\subsection{Stan Code}
```{r stan1, echo=T, eval=F}
data {
  int<lower=0> Npreds;         # number of covariates, including intercept
  int<lower=0> Npreds2;        # number of random effect covariates
  int<lower=0> Nplots;         # number of plots
  int<lower=0> Ntreats;        # number of treatments
  int<lower=0> Nobs;           # number of observations
  int<lower=0> Nyears;         # number of years
  vector[Nobs] y;              # vector of observations
  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
  int plot_id[Nobs];           # vector of plot ids
  int treat_id[Nobs];          # vector of treatment ids
  int year_id[Nobs];           # vector of year ids
}

parameters {
	vector[Npreds] beta;	             # overall coefficients
	vector[Nyears] year_off;           # vector of year effects
	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
	real<lower=0> sd_y;                # treatment-level observation std. dev.
	real<lower=0> sigma_year;          # year std. dev. hyperprior
	real<lower=0> phi;                 # dispersion parameter
}

transformed parameters {
  vector[Nobs] A;             # parameter for beta distn
  vector[Nobs] B;             # parameter for beta distn
  vector[Nobs] yhat;               # vector of expected values
  vector[Npreds2] u[Nplots];       # transformed plot random effects
  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix

  Sigma_u = diag_pre_multiply(sigma_u, L_u); # cholesky factor for plot-level covariance matrix 
  for(j in 1:Nplots)
    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes

  # regression model for expected values (one for each plot-year)
  for (i in 1:Nobs)
    yhat[i] = inv_logit(x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]);

  A = yhat * phi;
  B = (1.0 - yhat) * phi;
}

model {
  ####  PRIORS
  phi ~ cauchy(0, 5); 
  sigma_year ~ cauchy(0,2.5);
  sd_y ~ cauchy(0,2.5);
  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
  beta ~ normal(0,5);	             # priors on treatment coefficients
  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
                                    # correlation between plot level treatment effects
  sigma_u ~ cauchy(0,2.5);

  for(i in 1:Nplots)
		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope

	####  LIKELIHOOD
  y ~ beta(A, B); # observations vary according to beta distribution
}

generated quantities{
  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
}

```