### 2018-01-18 manuscript/Appendix2.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A2-\arabic{figure}}
\renewcommand\theequation{A2.\arabic{equation}}
\renewcommand\thetable{A2-\arabic{table}}  
\renewcommand\thesection{Section A2.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 2}} \\
A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

\section{Details on SOILWAT predictions}

We used a version of the SOILWAT soil moisture model [@Sala1992a] that has been developed specifically for use in semi-arid shrubland ecosystems [@Bradford2014].
SOILWAT uses daily weather data, ecosystem specific vegetation data, and site specific soil properties to estimate water balance processes.
Specifically, SOILWAT uses daily rainfall data to estimate rainfall interception by plants, evaporation of intercepted water, snow melt and redistribution, infiltration into the soil, percolation through the soil, evaporation from bare soil, transpiration from each soil layer, and drainage.
We parameterized SOILWAT using the generic sagebrush steppe parameters and local soil data [@Kleinhesselink2017b].
SOILWAT was forced by weather data collected at the US Sheep Experimental Station over the course of our experiment.

SOILWAT generates soil moisture predictions at several soil depths.
We averaged the daily predictions from the upper 40 cm of soil.
These predictions represent ambient conditions, similar to our control plots.
To generate soil moisture data for our treatment plots, we applied the statistical model described in the main text, which was also used to estimate treatment conditions from control conditions.
The time series of those predictions, along with our observations and statistical estimates, is shown in Figure A2-1.

```{r vwc, fig.cap="Time series of volumetric water content from March to June in each year from the observed measurements (solid lines), statistical estimates (dashed line), and SOILWAT (dotted line).", message=FALSE, echo=FALSE, comment=FALSE, warning=FALSE}

####
####  LOAD LIBRARIES -----------------------------------------------------------
####
library(tidyverse) # Data science packages
library(dplyr)     # Data summarizing and manipulating
library(stringr)   # Working with strings
library(ggthemes)  # Pleasing themes for ggplot

####
####  PLOT SOIL WATER BY TREATMENT ---------------------------------------------
####
mycols <- c("#009E73", "#D55E00", "#0072B2")

soilwat <- readRDS("../data/soil_moisture_data/SOILWAT_treatment_years.RDS") %>%
  select(date, year, Treatment, VWC_raw, doy) %>%
  rename(julian_date = doy) %>%
  mutate(julian_date = as.integer(julian_date))

soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_moisture.csv") %>%
  select(-year) %>%
  separate(simple_date, c("year","month","day")) %>%
  mutate(year = as.integer(year),
         Treatment = as.character(Treatment)) %>%
  ungroup() %>%
  left_join(soilwat, by = c("year","Treatment","julian_date")) %>%
  mutate(VWC_combo = ifelse(is.na(VWC)==TRUE, VWC_raw, VWC),
         VWC_source = ifelse(is.na(VWC)==TRUE, "soilwat", "observed")) %>%
  filter(month %in% c("03","04","05","06"))

ggplot(filter(soil_moisture, type == "observed"), aes(julian_date, VWC, group=Treatment, color=Treatment))+
  geom_line(data = soil_moisture, 
            aes(x=julian_date, y=VWC_raw, group=Treatment, color=Treatment),
            size=0.3, linetype=3)+
   geom_line(data = filter(soil_moisture, type == "predicted"), 
            aes(x=julian_date, y=VWC, group=Treatment, color=Treatment),
            size=0.3, linetype=2)+
  geom_line(size=0.3)+
  scale_color_manual(values = mycols, name="Treatment")+
  ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+
  xlab("Julian Day")+
  scale_y_continuous(breaks=seq(0,24,8))+
  scale_x_continuous(breaks=seq(0,250,20))+
  facet_grid(year~.)+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
        strip.background = element_blank(), 
        strip.text = element_text(size=6))

```

# References